msprime 1.0: an efficient and extensible coalescent simulation framework

### 2021-07-08 paper.tex
Partial draft of data interchange section

\subsection*{Integration with forwards-time simulations}
Forward simulations can also use tree
sequences~\citep{kelleher2018efficient,haller2018tree}. In particular,
SLiM 3.0~\citep{haller2019slim} and fwdpp~\citep{thornton2014cpp} can now output tree sequences.
Msprime can take tree sequences simulated by these forward simulators
as input and complete the simulation (``recapitation'').

@@

\subsection*{Data model}
The output of \msprime\ is a ``succinct tree sequence`` (or tree sequence,
for short) using the \tskit\ library. [Quick overview of the output formats
of other simulators, and the perf advantages of tree sequences.]

[Review the data model in Phastsim~\cite{demaio2021phastsim} and
Usher~\citep{turakhia2021ultrafast}.]

>>>>

@@

\subsection*{Data interchange and interoperability}
The point of simulations is to generate data and the efficiency with
which simulated data can be incorporated into an overall application
is an important---if often overlooked---issue.
For example, Approximate Bayesian Computation
(ABC)~\citep{beaumont2002approximate,csillery2010approximate}
and more recent machine learning
methods~\citep{sheehan2016deep,schrider2018supervised,flagel2019unreasonable}
perform population genetic inference by using large numbers of
simulations as training data. Clearly, simulation efficiency is
crucial since the size and number of simulations that can be performed
largely determines the scope of what can be inferred. Equally important,
however, is the efficiency with which the simulation results can be
transformed into the specific input required by the inference method.
In the case of ABC, this is usually a set of summary statistics of the
sequence data. General approaches such as
ABCtoolbox~\citep{wegmann2010abctoolbox} connect to
simulation and summary statistic computation programs via text streams
and are quite flexible, but the overhead involved in processing
the text is often a computational bottleneck.
Other methods such as DIYABC~\citep{cornuet2008inferring}
or PopABC~\citep{lopes2009popabc} avoid this bottleneck by
developing their own specialised simulators, tightly integrating them
with the inference method. While this is certainly more efficient,
there is a great deal of duplication of effort and significant
room for error.

\jkcomment{Stuff to cover}
\begin{itemize}

\item The core problem is that the field lacks an efficient mechanism for
    interchanging simulated data, and so developers are forced to choose
    between efficiency and duplication of effort.

\item What are the output formats of simulators? Newick, or sequence data
    basically. What's the problem with this? We don't have mutations.
    Much more efficient to have the tree and mutations than the
    sequence data. Fundamentally, the output of coalescent simulations
    should be trees, otherwise we are throwing a lot of information
    away.

\item Side note somewhere: with efficient access to branch length
    statisics~\citep{ralph2019efficiently}, maybe we don't need to
    simulate mutations at all.

\item Msprime uses the \tskit\ library~\citep{tskit2021tskit}, which provides
    extremely efficient interchange and data processing. A quick example
    of how efficient.

\item Tskit has facilited unprecedented interoperability between simulators:
    Forward simulations can also use tree
    sequences~\citep{kelleher2018efficient,haller2018tree}. In particular, SLiM
    3.0~\citep{haller2019slim} and fwdpp~\citep{thornton2014cpp} can now output
    tree sequences. Msprime can take tree sequences simulated by these forward
    simulators as input and complete the simulation (``recapitation'').
    See \cite{haller2018tree} for more details.

\item Recently, Phastsim/USHER~\cite{demaio2021phastsim,turakhia2021ultrafast}
    have used this idea of storing the
    tree plus mutations, but approach is crude compared to tskit.

\end{itemize}
